---
title: "GARCH_KF"
output: html_document
date: "2025-03-10"
---



```{r}
# install.packages("MASS")
# install.packages("ggplot")
library(MASS)
library(ggplot2)
library(tibble)
library(rugarch)
```

# Functions


```{r}
basis <- function(x, degree, i, knots){
  if(degree == 0){
    B <- ifelse((x>=knots[i])&(x<knots[i+1]),1,0)
  } else {
    if((knots[degree + i] - knots[i]) == 0){
      if(x != knots[i+degree]){
        alpha1 <- 0
      } else {
        return(1)
      }
    } else {
      alpha1 <- (x-knots[i])/(knots[degree+i] - knots[i])
    }
    if((knots[i+degree+1] - knots[i+1]) == 0){
      if(x != knots[i+degree]){
        alpha2 <- 0
      } else {
        return(1)
      }
    } else {
      alpha2 <- (knots[i+degree+1] - x) / (knots[i+degree+1] - knots[i+1])
    }
    B <- alpha1 * basis(x, (degree-1), i, knots) + 
      alpha2*basis(x, (degree-1), (i+1), knots)
  }
  return(B)
}
```

Construct the matrix $\mathbf{B}$, column by column. Matrix $B$ contains the basis functions computed through the Cox-deBoor recursive formulas using the function defined in the codechunk above, evaluated at different times to maturity (rows) and at different knot intervals (columns). For more details, check FURM on canvas.

```{r}
matrix_b <- function(x, degree=3, int_knots) { 
  # the x argument takes in a vector of time values that 
  # will be used to evaluate a design matrix of basis functions 
  # the degree argument specifies the highest degree of polynomials for
  # the basis functions
  # the int_knots argument takes in a vector of knots that will be used 
  # to determine the intervals of the piecewise function
  bound_knots <- int_knots[c(1, length(int_knots))] # this line creates bound knots
  knots <- c(rep(bound_knots[1], (degree+1)), int_knots[c(-1, -length(int_knots))], rep(bound_knots[2], (degree+1)))
  # the line above adds a couple of extra knots to each end of the int_knots vector because of the Cox-deBoor recursion
  K <- length(int_knots) + degree - 1 # number of columns in the Basis matrix
  B.mat <- matrix(0,nrow = length(x), ncol = K) # initialize the matrix
  for(j in 1:K) {
    B.mat[,j] <- sapply(X = x, FUN = basis, degree = degree, i = j, knots = knots) # add each column, one by one
  }
  return(B.mat) # return the matrix
}
```


```{r}
interp_yc <- function(yield_list, int_knots, degree = 3, d, last_tenor){
  # yield_list: parameter of the form of a list of data frames containing ZCB spot rate
  # int knots: the interior knots used for b-spline construction
  # degree: highest degree of polynomials for the basis functions
  # d: the date chosen to interpolate from the list
  # last_tenor: last tenor to interpolate in a day
  yield_list[[d]] <- data.frame(Maturity = yield_list[[d]]$Maturity,
                                ZERO_YLD1 = yield_list[[d]]$ZERO_YLD1)
  yc_df_pre <- rbind(data.frame(Maturity = 0, ZERO_YLD1 = 0), na.omit(yield_list[[d]]))
  last_row <- which(round(yc_df_pre$Maturity,3) == last_tenor)
  yc_df <- yc_df_pre[1:last_row,]
  yields <- c(0, yc_df$ZERO_YLD1)
  maturities <- c(0, as.numeric(yc_df$Maturity))
  x <- as.numeric(maturities) # maturity dates
  B <- matrix_b(x, degree=degree, int_knots = int_knots) 
  B_t_B <- t(B) %*% B
  # B is the design matrix on which the least squares coefficients will be calculated
  
  alphas <- solve(B_t_B) %*% t(B) %*% yields # OLS Formula for coefficients
  x2 <- seq(1/12, last_tenor, 1/12) # this range is used to simulate a continuous yield curve
  B2 <- matrix_b(x2, degree = degree, int_knots = int_knots) 
  # B2 is the matrix of basis functions but evaluated at a 'continuous' time (not really but close enough)
  
  interpolated_yields <- data.frame(Maturity = x2, ZERO_YLD1 = B2 %*% alphas) # create dataframes for plotting
  og_yields <- data.frame(ttm = maturities, yield = yields)

  return(interpolated_yields)
}
```

Interpolate a list of data with different dates.

```{r}
interpolate_list <- function(yield_list, start, T_, degree = 3){
  # yield_list: Parameter of the form of a list of data frames containing ZCB spot rate
  # start: starting date from the yield_list list
  # T_: length of time window
  # degree: highest degree of polynomials for the basis functions
  interpolated_yc <- list()
  k <- 1
  for(i in start:(start + T_ - 1)){
    lt_max <- max(yield_list[[i]]$Maturity) # This line of code basically chops all yields beyond 20
    avail_ylds <- na.omit(yield_list[[i]]$ZERO_YLD1)
    maturities <- yield_list[[i]]$Maturity
    N <- length(avail_ylds)
    if(N %in% c(5, 6)){
      int_knots <- c(0, quantile(maturities, probs = c(0, 0.5, 1)))
    } else if(N %in% c(7,8,9)){
      int_knots <- c(0,quantile(maturities, probs = c(0, 0.33, 0.66, 1)))
    } else if(N %in% 10:15){
      int_knots <- c(0,quantile(maturities, probs = c(0, 0.25, 0.5, .75, 1)))
    } else {
      int_knots <- c(0,quantile(maturities, probs = c(0, 0.20, 0.4, .6, .8, 1)))
    }
    interpolated_yc[[k]] <- interp_yc(yield_list = yield_list,
                                      int_knots = int_knots,
                                      d = i,
                                      last_tenor = lt_max,
                                      degree = degree)[3:240,]
    k <- k + 1
  }
  return(interpolated_yc)
}
```

Kalman Filter basis calculation 
```{r}
NS_loading <- function(lambda, mats) {
  loading <- rbind(rep(1, length(mats)),
                   (1 - exp(-lambda*mats)) / (lambda*mats),
                   (1 - exp(-lambda*mats)) / (lambda*mats) - exp(-lambda*mats))
  return(loading)
}
```


Standard Kalman Filter
```{r}
KF_NS <- function(par, yt, mats, lam) {
  # Standard Kalman Filter for Nelson Siegel model
  # Model:
  #   y_t = d_t + F_t' x_t + f_t + v_t, v_t ~ N(0, V), observation equation
  #   x_t = c + G x_{t-1} + w_t, w_t ~ N(0, W), state equation
  #   f_t = b*t + beta*cos(2*pi*t*dt) + eta*sin(2*pi*t*dt), seasonal effect
  # Inputs:
  #   par: a vector of parameters
  #   yt: the logarihm of futures prices
  #   T: maturities
  #   delivery_time: a vector of date, which is necessary if seasonality is "Constant"
  #   dt: delta t
  #   smoothing: a boolean variable indicate if Kalman Smoothing is required
  #   seasonality: "Constant" or "None"
  # Outputs:
  #   nll: the negative log likelihood
  #   ll_table: a vector to store cumulative log-likelihood at each time point - used to calculate Sandwich variance
  #   table_at_filter: a nT*2 matrix gives the filtered values of state variables.
  #   table_at_prediction: a (nT+1)*2 matrix gives the predicted values of state variables.
  #   table_at_smoother: a nT*2 matrix gives the smoothed values of state variables. The algorithm of Kalman Smoother is given by Bierman (1973) and De Jong (1989).
  #   ft: seasonal effect

 # require(lubridate)

  n_obs <- dim(yt)[1]
  n_contract <- dim(yt)[2]

  table_xt_filter <- matrix(0, nrow = n_obs, ncol = 3) # a_t|t
  table_Pt_filter <- array(0, dim = c(3, 3, n_obs)) # P_t|t
  table_xt_prediction <- matrix(0, nrow = n_obs+1, ncol = 3) # a_t|t-1
  table_Pt_prediction <- array(0, dim = c(3, 3, n_obs+1)) # P_t|t-1
  
  nll <- 0 # negative log-likelihood
  ll_table <- matrix(0, nrow = 1, ncol = n_obs) # table of log-likelihood

  table_et <- matrix(0, nrow = n_obs, ncol = n_contract) # e_t
  table_L <- array(0, dim = c(n_contract, n_contract, n_obs)) # Covariance of y
  table_y <- matrix(0, nrow = n_obs, ncol = n_contract) # y_hat

  # Parameters
  if (length(par) != 9+n_contract) {
    stop("Incorrect number of parameters. ")
  }

  phi0 <- c(par[1], par[2], par[3])
  phi1 <- diag(c(par[4], par[5], par[6]))
  x0 <- phi0 / (1 - diag(phi1))
  #lam <- par[7]
  sigma_w <- diag( par[7:9] )
  sigma_v <- diag( par[10: length(par)] )
  
  # Initialization
  xt_filter <- x0
  Pt_filter <- diag( diag(sigma_w) / (1 - diag(phi1)^2) )

  # Kalman Filter
  for (i in 1:n_obs) {
    lambda <- rbind(rep(1, n_contract),
                    (1 - exp(-lam * mats[i,])) / (lam * mats[i,]),
                    (1 - exp(-lam * mats[i,])) / (lam * mats[i,]) - exp(-lam * mats[i,]))

    # Prediction step
    xt_prediction  <- phi0 + phi1 %*% xt_filter # a_t+1|t
    Pt_prediction <- phi1 %*% Pt_filter %*% t(phi1) + sigma_w # P_t+1|t
    y_prediction <- t(lambda) %*% xt_prediction # ytilde_t|t-1 = d_t + F_t a_t|t-1

    # Filter step
    et <- yt[i, ] - t(y_prediction) # e_t = y_t - ytilde_t|t-1
    L <- t(lambda) %*% Pt_prediction %*% lambda + sigma_v # Covariance matrix of et
    invL <- solve(L) # inverse of L
    K <- Pt_prediction %*% lambda %*% invL # Kalman gain matrix: K_t

    xt_filter <- xt_prediction + K %*% t(et) # a_t
    Pt_filter <- (diag(3) - K %*% t(lambda)) %*% Pt_prediction # P_t
    #Pt_filter <- (diag(3) - K %*% t(lambda)) %*% Pt_prediction %*% t(diag(3) - K %*% t(lambda)) + K %*% sigma_v %*% t(K)

    # Update tables
    table_xt_filter[i, ] <- t(xt_filter)
    table_Pt_filter[, , i] <- Pt_filter
    table_xt_prediction[i+1, ] <- t(xt_prediction)
    table_Pt_prediction[, , i+1] <- Pt_prediction
    table_et[i, ] <- et
    table_L[, , i] <- L
    table_y[i, ] <- y_prediction

    if (det(L)<0) {
      message(i)
      message("matrix is not semi positive definite (KF)")
    }

    # Update likelihood
    nll <- nll + 0.5*log(det(L)) + 0.5*et %*% solve(L) %*% t(et)
    ll_table[i] <- -(0.5*length(yt[i, ])*log(2*pi) + 0.5*log(det(L)) + 0.5*et %*% solve(L) %*% t(et))
  }

  return(list(nll = nll,
              ll_table = ll_table,
              xt_filter = table_xt_filter,
              Pt_filter = table_Pt_filter,
              xt_prediction = table_xt_prediction,
              cov_y = table_L,
              y_hat = table_y))
}


```


Kalman Filter with GARCH
```{r}
KF_NS_GARCH <- function(par, yt, mats, lam, volatility_array) {
  # Standard Kalman Filter for Nelson Siegel model
  # Model:
  #   y_t = d_t + F_t' x_t + f_t + v_t, v_t ~ N(0, V), observation equation
  #   x_t = c + G x_{t-1} + w_t, w_t ~ N(0, W), state equation
  #   f_t = b*t + beta*cos(2*pi*t*dt) + eta*sin(2*pi*t*dt), seasonal effect
  # Inputs:
  #   par: a vector of parameters
  #   yt: the logarihm of futures prices
  #   T: maturities
  #   delivery_time: a vector of date, which is necessary if seasonality is "Constant"
  #   dt: delta t
  #   smoothing: a boolean variable indicate if Kalman Smoothing is required
  #   seasonality: "Constant" or "None"
  #   volatility_array: the array of Q_t covariance matrix fitted by rugarch ( input the array for simplicity, we could also use the GARCH paramter for for state covariance update)
  # Outputs:
  #   nll: the negative log likelihood
  #   ll_table: a vector to store cumulative log-likelihood at each time point - used to calculate Sandwich variance
  #   table_at_filter: a nT*2 matrix gives the filtered values of state variables.
  #   table_at_prediction: a (nT+1)*2 matrix gives the predicted values of state variables.
  #   table_at_smoother: a nT*2 matrix gives the smoothed values of state variables. The algorithm of Kalman Smoother is given by Bierman (1973) and De Jong (1989).
  #   ft: seasonal effect
  
 # require(lubridate)

  n_obs <- dim(yt)[1]
  n_contract <- dim(yt)[2]

  table_xt_filter <- matrix(0, nrow = n_obs, ncol = 3) # a_t|t
  table_Pt_filter <- array(0, dim = c(3, 3, n_obs)) # P_t|t
  table_xt_prediction <- matrix(0, nrow = n_obs+1, ncol = 3) # a_t|t-1
  table_Pt_prediction <- array(0, dim = c(3, 3, n_obs+1)) # P_t|t-1

  nll <- 0 # negative log-likelihood
  ll_table <- matrix(0, nrow = 1, ncol = n_obs) # table of log-likelihood

  table_et <- matrix(0, nrow = n_obs, ncol = n_contract) # e_t
  table_L <- array(0, dim = c(n_contract, n_contract, n_obs)) # Covariance of y
  table_y <- matrix(0, nrow = n_obs, ncol = n_contract) # y_hat

  # Parameters
  if (length(par) != 9+n_contract) {
    stop("Incorrect number of parameters. ")
  }

  phi0 <- c(par[1], par[2], par[3])
  phi1 <- diag(c(par[4], par[5], par[6]))
  x0 <- phi0 / (1 - diag(phi1))
  #lam <- par[7]
  sigma_w <- diag( par[7:9] )
  sigma_v <- diag( par[10: length(par)] )

  # Initialization
  xt_filter <- x0
  Pt_filter <- diag( diag(sigma_w) / (1 - diag(phi1)^2) )

  # Kalman Filter
  for (i in 1:n_obs) {
    lambda <- rbind(rep(1, n_contract),
                    (1 - exp(-lam * mats[i,])) / (lam * mats[i,]),
                    (1 - exp(-lam * mats[i,])) / (lam * mats[i,]) - exp(-lam * mats[i,]))
    sigma_w <- volatility_array[,,i]
    
    # Prediction step
    xt_prediction  <- phi0 + phi1 %*% xt_filter # a_t+1|t
    Pt_prediction <- phi1 %*% Pt_filter %*% t(phi1) + sigma_w # P_t+1|t
    y_prediction <- t(lambda) %*% xt_prediction # ytilde_t|t-1 = d_t + F_t a_t|t-1
    
    # Filter step
    
    et <- yt[i, ] - t(y_prediction) # e_t = y_t - ytilde_t|t-1
    L <- t(lambda) %*% Pt_prediction %*% lambda + sigma_v # Covariance matrix of et
    invL <- solve(L) # inverse of L
    K <- Pt_prediction %*% lambda %*% invL # Kalman gain matrix: K_t
    
    xt_filter <- xt_prediction + K %*% t(et) # a_t
    Pt_filter <- (diag(3) - K %*% t(lambda)) %*% Pt_prediction # P_t
    #Pt_filter <- (diag(3) - K %*% t(lambda)) %*% Pt_prediction %*% t(diag(3) - K %*% t(lambda)) + K %*% sigma_v %*% t(K)

    # Update tables
    table_xt_filter[i, ] <- t(xt_filter)
    table_Pt_filter[, , i] <- Pt_filter
    table_xt_prediction[i+1, ] <- t(xt_prediction)
    table_Pt_prediction[, , i+1] <- Pt_prediction
    table_et[i, ] <- et
    table_L[, , i] <- L
    table_y[i, ] <- y_prediction


    if (det(L)<0) {
      message(i)
      message("matrix is not semi positive definite (KF)")
    }

    # Update likelihood
    nll <- nll + 0.5*log(det(L)) + 0.5*et %*% solve(L) %*% t(et)
    ll_table[i] <- -(0.5*length(yt[i, ])*log(2*pi) + 0.5*log(det(L)) + 0.5*et %*% solve(L) %*% t(et))
  }
  
  

  return(list(nll = nll,
              ll_table = ll_table,
              xt_filter = table_xt_filter,
              Pt_filter = table_Pt_filter,
              xt_prediction = table_xt_prediction,
              cov_y = table_L,
              y_hat = table_y))
}



```


Yule-Walker Estimate for AR(1)
```{r}
yule_walker_ar1 <- function(ts) {
  # Estimates the parameters of an AR(1) process using the Yule-Walker equations.
  #
  # Parameters:
  # ts: A univariate time series (numeric vector).
  #
  # Returns:
  # A list containing:
  #   phi: The estimated autoregressive coefficient.
  #   sigma2: The estimated variance of the innovations.
  #   c: The estimated intercept term.

  # Compute the sample mean (used for centering the time series)
  mu <- mean(ts)
  
  # Center the time series
  ts_centered <- ts - mu
  
  # Compute the autocovariance at lag 0 and lag 1
  gamma0 <- var(ts_centered) * (length(ts_centered) - 1) / length(ts_centered)
  gamma1 <- sum(ts_centered[-1] * ts_centered[-length(ts_centered)]) / length(ts_centered)
  
  # Estimate phi using the Yule-Walker equation
  phi <- gamma1 / gamma0
  
  # Estimate the intercept c
  c <- mu * (1 - phi)
  
  # Estimate the innovation variance sigma2
  sigma2 <- gamma0 * (1 - phi^2)
  
  # Return the estimates
  return(list(phi = phi, sigma2 = sigma2, c = c))
}

```

OLS for Nelson Siegel
```{r}
fit_nelson_siegel <- function(yield_list, lambda, start, tenors, T_){
  # yield_list: Parameter of the form of a list of data frames containing ZCB spot rate
  # lambda: Individual lambda parameter for NS
  # start: starting date from the yield_list list
  # tenors: list of time to maturities
  # T_: length of time window

  indices <- which(round(yield_list[[start]]$Maturity, 2) %in% round(tenors, 2))
  maturities <- yield_list[[1]]$Maturity[indices]
  N <- length(maturities)
  
  term1 <- 1
  term2 <- (1 - exp(-maturities*lambda)) / (maturities*lambda)
  term3 <- ((1 - exp(-maturities*lambda)) / (maturities*lambda)) - exp(-maturities*lambda)
  Phi <- cbind(term1, term2, term3) # Construct Phi matrix for NS
  
  Y_mat <- matrix(0, nrow = N, # matrix of N by T, containing yields for each tenor (columns)
                  ncol = T_) # where each column represents a different date
  j <- 1
  for(t in start:(start + T_- 1)){
    Y_mat[,j] <- yield_list[[t]]$ZERO_YLD1[indices]
    j <- j + 1
    phitphi_1phit <- solve(t(Phi) %*% Phi, t(Phi)) # OLS for the coefficients for every single day
    betas_t <- phitphi_1phit %*% Y_mat  
    betas <- rowSums(betas_t) / T_ # average all coefficients
    eps <- matrix(0, nrow = N, ncol = T_) # matrix of errors for each day (column) and each tenor (row)
    for(t in 1:T_){
      eps[,t] <- Y_mat[,t] - Phi %*% betas # Populate errors
    }
    sig_hat2 <- sum(as.vector(eps)^2) / (N * T_ - 3) # take mean squared error (MLE Estimator)
  }

  return(list(betas = betas, # fitted betas static: 1*3   dynamic: T*3
              sigma2 = sig_hat2, # MSE
              lambda = lambda, # lambda(input)
              cov_mat_betas = sig_hat2 * solve(t(Phi) %*% Phi), # Nelson Siegel design matrix N*3
              eps = eps,
              Phi = Phi)) # residuals N*T
  
}
```








# Data Processing

Import data and we only take the last 300 days
```{r}
clean_yield_raw <-read.csv("./Clean_Yields.csv")
maturity_clean <- as.vector(c(1/12, 1/4, 1/2, 1, 2, 3, 5, 7, 10, 20, 30))
clean_yield_raw <- na.omit(as.matrix(clean_yield_raw))
st_louis_Fed_yields <- list()

for(i in 1:nrow(clean_yield_raw)){
  one_date_matrix <- tibble(
    Maturity = maturity_clean,
    ZERO_YLD1 = as.numeric(clean_yield_raw[i,2:12 ]),
    START_DT = format(as.Date(clean_yield_raw[i,1], format="%Y-%m-%d"), "%Y%m%d")
  )
  one_date_matrix <- na.omit(one_date_matrix)

  st_louis_Fed_yields[[i]] <- one_date_matrix
}

head(st_louis_Fed_yields)
```

Interpolate Data
```{r}
data_list_300 <- st_louis_Fed_yields[c((length(st_louis_Fed_yields) - 299):length(st_louis_Fed_yields))]
interp252 <- interpolate_list(data_list_300, 1, 300) # only take the first 100 dates

```

Prepared the interpolated data to matrix form.
```{r}
tenors = c(1/12, 5, 10, 15, 20)
n_matrices <- length(interp252)
n_tenors <- length(tenors)

training_data <- matrix(NA, nrow = n_matrices, ncol = n_tenors)
colnames(training_data) <- tenors  

# Iterate through each matrix in the list
for (i in 1:n_matrices) {
  mat <- interp252[[i]] 

# Match tenors to the closest Maturity column values
# Find the index of the closest maturity (I did this beaause somtimes 0.8333 ! = 1/12 in the interpolate data)
matched_values <- sapply(tenors, function(tenor) {
  rounded_maturity <- round(mat[, 1], 2) 
  rounded_tenor <- round(tenor, 2)

  idx <- which.min(abs(rounded_maturity - rounded_tenor))
  
  if (length(idx) > 0) {
    return(mat[idx, 2]) 
  } else {
    return(NA) 
  }
})
  
  training_data[i, ] <- matched_values  # Store extracted values
}


head(training_data)

```











# Paramter Estimation and Model Fitting 


Fit the latent state paramter (L, S, C) by OLS
```{r}
tsBetas <- matrix(NA, nrow = 3, ncol = 300) # state parameter estimate using OLC
obs_res <- matrix(NA, nrow = , ncol = 300) # residual of observation
for(i in 1:300) {
  ols_fit <- fit_nelson_siegel(interp252, 0.1, tenors = c(1/12, 5, 10, 15, 20), start = i, T_ = 1)
  tsBetas[,i] <- ols_fit$betas
  obs_res[,i] <- ols_fit$sigma2
}
```


The state transition matri A, the output are 3 diagonal term.
```{r}
AR1_level <- yule_walker_ar1(tsBetas[1,])
AR1_slope <- yule_walker_ar1(tsBetas[2,])
AR1_curv <- yule_walker_ar1(tsBetas[3,])

AR1_level
AR1_slope
AR1_curv
```





Fit a standard KF using the paramters A, Q, R.
```{r}
sigOLS2 <- mean(obs_res) # R: covariance matrix of observation noise (assumed to be diagonal)
tenors <- c(1/12, 5, 10, 15, 20) # tenor of each time
mats_matrix_real <- matrix(rep(tenors, each = 300), nrow = 300, byrow = TRUE)


parameters <- c(AR1_level$c, AR1_slope$c, AR1_curv$c, AR1_level$phi, AR1_slope$phi, AR1_curv$phi, AR1_level$sigma2, AR1_slope$sigma2,
            AR1_curv$sigma2, rep(sigOLS2, length(tenors)))

real_result<- KF_NS(parameters, training_data, mats_matrix_real, 0.1)
head(real_result$y_hat) # kalman filter prediction for yield data

```


Use the standard Kalman Filter get the predicted state estimated and the filtered state. We get the residual for GARCH(1,1) model fitting.
```{r}
filtered_x <- real_result$xt_filter
predicted_x <- real_result$xt_prediction[-301,]


Beta_res <- filtered_x - predicted_x

```

Use the state residual to fit the GARCH(1,1) model.
```{r}
garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "norm"  # Normal distribution for residuals
)

garch_beta_1 <- ugarchfit(spec = garch_spec, data = Beta_res[,1])
garch_beta_2 <- ugarchfit(spec = garch_spec, data = Beta_res[,2])
garch_beta_3 <- ugarchfit(spec = garch_spec, data = Beta_res[,3])

garch_beta_1
garch_beta_2
garch_beta_3


beta_1_fit <- garch_beta_1@fit$coef
beta_2_fit <- garch_beta_2@fit$coef
beta_3_fit <- garch_beta_3@fit$coef


vol_param <- rbind(beta_1_fit[2:4], beta_2_fit[2:4], beta_3_fit[2:4])
vol_param # parameters for volatility
```

Get the predicted volatility in to data frame.
```{r}
# Extract conditional volatility for each time t
volatility_beta_1 <- sigma(garch_beta_1)
volatility_beta_2 <- sigma(garch_beta_2)
volatility_beta_3 <- sigma(garch_beta_3)

# Combine into a data frame
volatility_df <- data.frame(
  Time = 1:length(volatility_beta_1),  # Assuming all series have the same length
  Volatility_Beta_1 = volatility_beta_1,
  Volatility_Beta_2 = volatility_beta_2,
  Volatility_Beta_3 = volatility_beta_3
)

# View the volatility time series
head(volatility_df)  # Show the first few rows

```


Convert the Q_t matrix over time into array, which is T number of 3*3 matrices. 
For inputnig Q_t at once to the function, this may not be a good choice for prediction, but it is efficient if we want to see the tracking.
For prediction, we need to input the volatility paramters $\omega$, $\alpha_1$, $\beta1$ for 1 step aheadprediction. 
```{r}


T_ <- length(volatility_beta_1)
volatility_array <- array(0, dim = c(3, 3, 300))
for (t in 1:300) {
  volatility_array[,,t] <- matrix( c(as.numeric(volatility_beta_1[t]),0,0,0, as.numeric(volatility_beta_2[t]),0,0,0, as.numeric(volatility_beta_3[t])), 3, 3 )
}


volatility_array[,,1:5]

```

Fit a KF-GARCH model.
```{r}
KF_G_res<-KF_NS_GARCH(parameters, training_data, mats_matrix_real, 0.1, volatility_array = volatility_array)
```







# Model result and Comparison


Compare the prediction values of the lastest 5 days and log-likelihhod function of the fit.
```{r}
tail(KF_G_res$y_hat)
tail(real_result$y_hat)


cat("Log Likelihood of Kalman Filter:", real_result$nll,"\n")
cat("Log Likelihood of Kalman Filter-Garch:", KF_G_res$nll)
```



Plot for standard KF
```{r}
# Extract y_hat (KF predictions) and real interpolated data
y_hat <- real_result$y_hat  # Predicted values from KF
y_real <- training_data  # Real interpolated data

# Extract covariance matrix of y for confidence intervals
y_cov <- real_result$cov_y  # Covariance matrices for each time step

# Compute 95% confidence interval
n_obs <- nrow(y_hat)
n_contract <- ncol(y_hat)

# Initialize matrices for upper and lower bounds
ci_upper <- matrix(NA, n_obs, n_contract)
ci_lower <- matrix(NA, n_obs, n_contract)

for (i in 1:n_obs) {
  std_err <- sqrt(diag(y_cov[,,i]))  # Extract standard deviation for each tenor
  ci_upper[i, ] <- y_hat[i, ] + 1.96 * std_err  # Upper bound
  ci_lower[i, ] <- y_hat[i, ] - 1.96 * std_err  # Lower bound
}

# Convert to a data frame for ggplot (taking the first tenor as an example)
df <- data.frame(
  Time = 1:n_obs, 
  Real = y_real[, 2],  # Taking the first tenor as an example
  Predicted = y_hat[, 2],
  CI_Upper = ci_upper[, 2],
  CI_Lower = ci_lower[, 2]
)

df <- data.frame(
  Time = 1:n_obs, 
  Real = y_real[, 3],  
  Predicted = y_hat[, 3],
  CI_Upper = ci_upper[, 3],
  CI_Lower = ci_lower[, 3]
)


# loop to genrate plot for slong 5 different maturities
for (i in 2:5) {
  df <- data.frame(
    Time = 1:n_obs, 
    Real = y_real[, i],  
    Predicted = y_hat[, i],
    CI_Upper = ci_upper[, i],
    CI_Lower = ci_lower[, i]
  )
  
  plot <- ggplot(df, aes(x = Time)) +
    geom_line(aes(y = Real, color = "Real Data"), size = 1) +
    geom_line(aes(y = Predicted, color = "Predicted Data"), size = 1, linetype = "dashed") +
    geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper), fill = "blue", alpha = 0.2) +  
    labs(title = paste("Kalman Filter Prediction vs Real Data - Tenor", i),
         y = paste("Yield (Tenor", i, ")"),
         x = "Time") +
    scale_color_manual(values = c("Real Data" = "black", "Predicted Data" = "red", "Predicted 95% Confidence Interval" = "blue")) +
    theme_minimal()
  
  print(plot)
}


```



Plot for KF-GARCH
```{r}
y_hat <- KF_G_res$y_hat  # Predicted values from KF
y_real <- training_data  # Real interpolated data

y_cov <- KF_G_res$cov_y

n_obs <- nrow(y_hat)
n_contract <- ncol(y_hat)

# Initialize matrices for upper and lower bounds
ci_upper <- matrix(NA, n_obs, n_contract)
ci_lower <- matrix(NA, n_obs, n_contract)

for (i in 1:n_obs) {
  std_err <- sqrt(diag(y_cov[,,i]))
  ci_upper[i, ] <- y_hat[i, ] + 1.96 * std_err  # Upper bound of KF-GARCH
  ci_lower[i, ] <- y_hat[i, ] - 1.96 * std_err  # Lower bound of KF-GARCH
}

tenors <- c("1/12 year", "5-Year", "10-Year", "20-Year", "30-Year")

# loop to genrate plot for slong 5 different maturities
for (i in 2:5) { 
  df <- data.frame(
    Time = 1:n_obs, 
    Real = y_real[, i],  
    Predicted = y_hat[, i],
    CI_Upper = ci_upper[, i],
    CI_Lower = ci_lower[, i]
  )

  plot <- ggplot(df, aes(x = Time)) +
    geom_line(aes(y = Real, color = "Real Data"), size = 1) +
    geom_line(aes(y = Predicted, color = "Predicted Data"), size = 1, linetype = "dashed") +
    geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper), fill = "blue", alpha = 0.2) +  
    labs(title = paste("GARCH Kalman Filter Prediction vs Real Data -", tenors[i]),
         y = paste("Yield (", tenors[i], ")"),
         x = "Time") +
    scale_color_manual(values = c("Real Data" = "black", "Predicted Data" = "red", "Predicted 95% Confidence Interval" = "blue")) +
    theme_minimal()
  
  print(plot) 
}
```


Comparison of KF and KF-GARCH yield rate prediction and true yield rate
```{r}

tenor <- c("1 Month", "5 Years", "10 Years", "20 Years", "30 Years")

# loop to genrate plot for slong 5 different maturities
for (tenor_index in 2:5) {
  
  df <- data.frame(
    Time = 1:n_obs, 
    Real = y_real[, tenor_index],  
    KF_Predicted = real_result$y_hat[, tenor_index],
    KF_GARCH_Predicted = KF_G_res$y_hat[, tenor_index]
  )
  
  # Confidence intervals for KF
  df$CI_Upper_KF <- df$KF_Predicted + 1.96 * sqrt(diag(real_result$cov_y[,,tenor_index]))
  df$CI_Lower_KF <- df$KF_Predicted - 1.96 * sqrt(diag(real_result$cov_y[,,tenor_index]))

  # Confidence intervals for KF-GARCH
  df$CI_Upper_KF_GARCH <- df$KF_GARCH_Predicted + 1.96 * sqrt(diag(KF_G_res$cov_y[,,tenor_index]))
  df$CI_Lower_KF_GARCH <- df$KF_GARCH_Predicted - 1.96 * sqrt(diag(KF_G_res$cov_y[,,tenor_index]))

  # Create the plot
  plot <- ggplot(df, aes(x = Time)) +
    geom_line(aes(y = Real, color = "Real Data"), size = 1) +
    geom_line(aes(y = KF_Predicted, color = "KF Prediction"), size = 1, linetype = "dashed") +
    geom_line(aes(y = KF_GARCH_Predicted, color = "KF-GARCH Prediction"), size = 1, linetype = "dotted") +
    geom_ribbon(aes(ymin = CI_Lower_KF, ymax = CI_Upper_KF, fill = "KF 95% CI"), alpha = 0.2) +
    geom_ribbon(aes(ymin = CI_Lower_KF_GARCH, ymax = CI_Upper_KF_GARCH, fill = "KF-GARCH 95% CI"), alpha = 0.2) +
    labs(
      title = paste("Kalman Filter vs KF-GARCH Prediction -", tenor[tenor_index], "bond"),
      y = "Yield",
      x = "Time"
    ) +
    scale_color_manual(values = c(
      "Real Data" = "black",
      "KF Prediction" = "blue",
      "KF-GARCH Prediction" = "green"
    )) +
    scale_fill_manual(values = c(
      "KF 95% CI" = "blue",
      "KF-GARCH 95% CI" = "green"
    )) +
    guides(
      fill = guide_legend(title = "Confidence Interval"),
      color = guide_legend(title = "Line")
    ) +
    theme_minimal()
  
  
  print(plot)
  # ggsave(filename = paste0("Prediction_", gsub(" ", "_", tenor[tenor_index]), ".png"),
  #         plot = plot, width = 8, height = 6, dpi = 300)
}


```



Comparison of KF and KF-GARCH state prediction & state update
```{r}
state_names <- c("Level", "Slope", "Curvature")

for (i in 1:3) {
  df_x <- data.frame(
    Time = 1:n_obs,
    KF_Predicted = real_result$xt_prediction[1:300, i],  # KF predicted x
    KF_Filtered = real_result$xt_filter[, i],       # KF filtered x
    KF_GARCH_Predicted = KF_G_res$xt_prediction[1:300, i],  # KF-GARCH predicted x
    KF_GARCH_Filtered = KF_G_res$xt_filter[, i],       # KF-GARCH filtered x
    CI_Upper_KF = real_result$xt_filter[, i] + 1.96 * sqrt(real_result$Pt_filter[i, i, ]),
    CI_Lower_KF = real_result$xt_filter[, i] - 1.96 * sqrt(real_result$Pt_filter[i, i, ]),
    CI_Upper_KF_GARCH = KF_G_res$xt_filter[, i] + 1.96 * sqrt(KF_G_res$Pt_filter[i, i, ]),
    CI_Lower_KF_GARCH = KF_G_res$xt_filter[, i] - 1.96 * sqrt(KF_G_res$Pt_filter[i, i, ])
  )

  # **1. Plot for Kalman Filter (KF) Only**
  p1 <- ggplot(df_x, aes(x = Time)) +
    geom_line(aes(y = KF_Filtered, color = "KF Filtered"), size = 1) +
    geom_line(aes(y = KF_Predicted, color = "KF Predicted"), size = 1, linetype = "dashed") +
    geom_ribbon(aes(ymin = CI_Lower_KF, ymax = CI_Upper_KF), fill = "blue", alpha = 0.2) + 
    labs(title = paste("Kalman Filter -", state_names[i]),
         y = state_names[i],
         x = "Time") +
    scale_color_manual(values = c("KF Filtered" = "red", "KF Predicted" = "black")) +
    theme_minimal()

  print(p1)  # Display KF plot

  # **2. Plot for KF-GARCH Only**
  p2 <- ggplot(df_x, aes(x = Time)) +
    geom_line(aes(y = KF_GARCH_Filtered, color = "KF-GARCH Filtered"), size = 1) +
    geom_line(aes(y = KF_GARCH_Predicted, color = "KF-GARCH Predicted"), size = 1, linetype = "dotted") +
    geom_ribbon(aes(ymin = CI_Lower_KF_GARCH, ymax = CI_Upper_KF_GARCH), fill = "green", alpha = 0.2) +  
    labs(title = paste("KF-GARCH -", state_names[i]),
         y = state_names[i],
         x = "Time") +
    scale_color_manual(values = c("KF-GARCH Filtered" = "blue", "KF-GARCH Predicted" = "green")) +
    theme_minimal()

  print(p2) 

  p3 <- ggplot(df_x, aes(x = Time)) +
    geom_line(aes(y = KF_Filtered, color = "KF Filtered"), size = 1) +
    geom_line(aes(y = KF_GARCH_Filtered, color = "KF-GARCH Filtered"), size = 1, linetype = "solid") +
    geom_ribbon(aes(ymin = CI_Lower_KF, ymax = CI_Upper_KF), fill = "blue", alpha = 0.2) +
    geom_ribbon(aes(ymin = CI_Lower_KF_GARCH, ymax = CI_Upper_KF_GARCH), fill = "green", alpha = 0.2) +
    labs(title = paste("Comparison: KF vs KF-GARCH -", state_names[i]),
         y = state_names[i],
         x = "Time") +
    scale_color_manual(values = c("KF Filtered" = "blue", "KF-GARCH Filtered" = "green")) +
    theme_minimal()

  print(p3)  # Display Comparison plot
}


```







